Enhancing Adversarial Robustness for Deep Metric Learning

### 2021-10-25 robdml.tex
tex

\cref{fig:hm}

by referencing
another benign (but harder) sample triplet.

make adversarially training a DML model
with triplet loss feasible without model collapse, so the model could
efficiently learn from a strong adversary.
%
Assume there are two data triplets, namely ``source triplet'' and ``destination
triplet'' with the same anchor example, but sampled using different triplet
sampling strategies.
%
And the ``destination triplet'' is expected to result in a higher loss value
than another.
%
HM aims at creating adversarial examples of the ``source triplet'', that would
increase the corresponding loss value until that of the unperturbed
``destination triplet''.
%
This can also be interpreted as maximizing the loss value but ``stops early''
at a certain point.
%
The resulting perturbed ``source triplet'' can be used for adversarial training
of the model.

Given an anchor image $X_a$, we sample a positive image $X_p^S$ (in the same
class as the anchor) and a negative image $X_n^S$ (in different class as the
anchor) with a certain sampling strategy (\eg, semi-hard~\cite{facenet}).
%
% For convenience we call this triplet as ``source triplet''.
%
Then its \emph{hardness} is defined as
%
\begin{equation}
%
H(X_a^S,X_p^S,X_n^S)=d_\phi(X_a,X_p^S)-d_\phi(X_a,X_n^S),
%
\end{equation}
%
which is an internal
part of the triplet loss.
%
For convenience we call this as ``source hardness'', denoted as $H_\mathsf{S}$.

@@

Then, we want change (increase) the hardness level from $H_\mathsf{S}$ into $H_\mathsf{D}$, by
finding adversarial examples for $(X_a, X_p^S, X_n^S)$.
%
Note, it is not expected to lower the hardness during the attack, namely
$E[H_\mathsf{D}]$ should be greater or equal to $E[H_\mathsf{S}]$.
%
We denote the the hardness of adversarial examples as $\tilde{H}_S=H(X_a{+}r_a,
X_p^S{+}r_p, X_n^S{+}r_n)$.
%
This is named as \emph{Hardness Manipulation} (HM), which is implemented as
follows:
%
\begin{equation}
	%
	\bar{r}_a, \bar{r}_p, \bar{r}_n = \argmin_{r_a,r_p,r_n} \big\|\min(0,
	\tilde{H}_\mathsf{S} - H_\mathsf{D}) \big\|_2^2.
	%
	\label{eq:hm}
	%
\end{equation}
%
The $\min(0,\cdot)$ part in \cref{eq:hm} truncates the gradient when $H_\mathsf{S}>H_\mathsf{D}$,
automatically stopping the optimization.
%
This is because $H_\mathsf{S}$ will be reduced once greater than $H_\mathsf{D}$ in standard Mean
Square Error.
%
The optimization problem can be solved by projected gradient
descent~\cite{madry}.
%
And the resulting adversarial examples are used for adversarially training the
DML model:
%
$L_\text{T}(\phi(X_a+\bar{r}_a), \phi(X_p^S+\bar{r}_p),
\phi(X_n^S+\bar{r}_n))$.

@@

\subsection{Gradual Adversary}
\label{sec:32}

>>>>

Given an image triplet ($\mA$, $\mP$, $\mN$) sampled with a certain sampling
strategy (\eg, semi-hard~\cite{facenet}) within a mini-batch, we define its
\emph{hardness} as a scalar:
%
\begin{equation}
%
H(\mA,\mP,\mN)=d_\phi(\mA,\mP)-d_\phi(\mA,\mN),
%
\end{equation}
%
which is an internal part of the triplet loss.
%
For convenience we call this triplet as ``\emph{source} triplet'', and the
corresponding $H(\cdots)$ value as ``\emph{source} hardness'', denoted as
$H_\mathsf{S}$.

@@

Then, we want change (increase) the hardness level from $H_\mathsf{S}$ into $H_\mathsf{D}$, by
finding adversarial examples for $(X_a, X_p^S, X_n^S)$.
%
Note, it is not expected to lower the hardness during the attack, namely
$E[H_\mathsf{D}]$ should be greater or equal to $E[H_\mathsf{S}]$.
%
We denote the the hardness of adversarial examples as $\tilde{H}_S=H(X_a{+}r_a,
X_p^S{+}r_p, X_n^S{+}r_n)$.
%
This is named as \emph{Hardness Manipulation} (HM), which is implemented as
follows:
%
\begin{equation}
	%
	\bar{r}_a, \bar{r}_p, \bar{r}_n = \argmin_{r_a,r_p,r_n} \big\|\min(0,
	\tilde{H}_\mathsf{S} - H_\mathsf{D}) \big\|_2^2.
	%
	\label{eq:hm}
	%
\end{equation}

\cref{fig:hm}
%
The $\min(0,\cdot)$ part in \cref{eq:hm} truncates the gradient when $H_\mathsf{S}>H_\mathsf{D}$,
automatically stopping the optimization.
%
This is because $H_\mathsf{S}$ will be reduced once greater than $H_\mathsf{D}$ in standard Mean
Square Error.
%
The optimization problem can be solved by projected gradient
descent~\cite{madry}.
%
And the resulting adversarial examples are used for adversarially training the
DML model:
%
$L_\text{T}(\phi(X_a+\bar{r}_a), \phi(X_p^S+\bar{r}_p),
\phi(X_n^S+\bar{r}_n))$.

@@

%
Assume there are two data triplets, namely ``source triplet'' and ``destination
triplet'' with the same anchor example, but sampled using different triplet
sampling strategies.
%
And the ``destination triplet'' is expected to result in a higher loss value
than another.
%
HM aims at creating adversarial examples of the ``source triplet'', that would
increase the corresponding loss value until that of the unperturbed
``destination triplet''.
%
This can also be interpreted as maximizing the loss value but ``stops early''
at a certain point.
%
The resulting perturbed ``source triplet'' can be used for adversarial training
of the model.
\subsection{Gradual Adversary}
\label{sec:32}
